### 2019-01-09 paper/sex_selection_spacing-ver2.tex
Add graphs on recall error and simply cut-off to 22 years or more.

NFHS enumerators probe for any missed births, although the method depends on the survey.
NFHS-1 probe for each calendar birth interval that is 4 or more years.
NFHS-2 asked for stillbirths, spontaneous and induced abortions and also probed 
for each calendar birth interval 4 or more years.
NFHS-3 and NFHS-4 did not directly use birth intervals, but asked whether were there any 
other live births between (name of previous birth) and (name), including any children who 
died after birth, and asked for births before the birth listed as first birth and
after the last birth listed as the last birth.

@@

The problem with the above approach is that the year of birth is affected by recall error; 
for example, a second born child listed as first-born is born later than the real 
first born child.
Recorded year of marriage should, however, be affected neither by parental recall error 
nor the use of sex-selective abortions.
Tables \ref{tab:recallMarriageBO1} and \ref{tab:recallMarriageBO2}, therefore, shows sex 
ratios of children recorded as first-born and second-born by year of parents' marriage, 
together with tests for whether the observed sex ratio is significantly higher than the 
natural sex ratio and whether more recent surveys show a higher sex ratio for the cohort 
than earlier surveys.

The basic recall error pattern remains with women married longer ago more
likely to report that their first-born is a boy.
If anything the pattern is more consistent when using year of marriage than 
year of birth. 
Before the introduction of sex selection there are only two instances  across the two 
parities where the observed sex ratio is lower for women married longer ago than for women 
married closer to the survey date.%
\footnote{
For first-borns it is between 1970--1974 and 1975-1979 for NFHS-3 and for second-borns
it is between 1975--1979 and 1980--1984 for NFHS-1.
}
Similarly, comparing women married in the same five-year period across surveys shows
that women married longer ago are more likely to report having a son.

@@

The relationship between length of marriage and recall error can also seen in 
Figure \ref{fig:sexRatioMarriage}, which shows the observed sex ratio for children reported
as first born as a function of duration of marriage at the time of the survey combining 
all four surveys.
The solid line is the sex ratio of children reported 
as first born by the number of years between the survey and marriage, 
the dashed lines indicate the 95 percent confidence interval 
and the horizontal line the natural sex ratio (approximately 0.512).
To ensure sufficient cell sizes the years are grouped in twos.
In line with the results from Tables \ref{tab:recallMarriageBO1} and 
\ref{tab:recallMarriageBO2}, the observed ratio of boys 
is increasingly above the expected value the longer ago the parents were married.

\begin{figure}[htp]%
\centering
\includegraphics[width=0.9\textwidth]{recall_sex_ratio_marriage}
\caption{Sex ratio of ``first'' births by duration of marriage}
\label{fig:sexRatioMarriage}
\end{figure}

The increasingly unequal sex ratio with increasing marriage duration suggests that
a solution to the recall error problem is to drop observations for 
women who were married ``too far'' from the survey year.
The main problem is establishing what the best cut-off point should be.
As Tables \ref{tab:recallMarriageBO1} and \ref{tab:recallMarriageBO2} shows there are 
differences in recall error across the three surveys and between parities.
One possibility is simply to drop all women married in cohorts where the observed
sex ratio is substantially higher than the natural sex ratio, but even this is not
straightforward.
For ``first'' born this would imply dropping all women married before 1980, but
for ``second'' born it would be imply dropping all married before 1970 for NFHS-1, before 1975 
for NFHS-2, and before 1980 NFHS-3.
Because of the differences between ``first'' and ``second'' born in the different surveys
I vary the cut-off point by survey round.

NFHS-3 is the most straightforward in that the implied cut-off point is the same for
both first and second born.
Based on Tables \ref{tab:recallMarriageBO1} and \ref{tab:recallMarriageBO2}, women in 
NFHS-3 who were married 26 years or
more before the survey date are dropped.
For NFHS-1 and NFHS-2 the trade-off is between ensuring enough information and not bias 
the results unduly.
Given that sex selective abortions are more likely in higher parity birth I base the
cut-off point on the second births instead of the first births.
Hence, for NFHS-1 women who were married 22 years or more before the survey date 
were dropped and for NFHS-2 women who were married 23 year or more before the survey date 
were dropped.

@@

>>>>

NFHS enumerators probe for any missed births, although the method depends on the survey.
NFHS-1 probe for each calendar birth interval that is four or more years.
NFHS-2 asked for stillbirths, spontaneous and induced abortions and also probed 
for each calendar birth interval four or more years.
NFHS-3 and NFHS-4 did not directly use birth intervals, but asked whether there were any 
other live births between (name of previous birth) and (name), including any children who 
died after birth, and asked for births before the birth listed as first birth and
after the last birth listed as the last birth.

@@

The problem with the above approach is that the year of birth is affected by recall error; 
a second born child listed as first-born is born later than the real first born child.
Year of marriage should, however, be affected neither by parental recall error 
nor the use of sex-selective abortions.
Tables \ref{tab:recallMarriageBO1} and \ref{tab:recallMarriageBO2}, therefore, shows sex 
ratios of children recorded as first-born and second-born by year of parents' marriage, 
together with tests for whether the observed sex ratio is significantly higher than the 
natural sex ratio and whether more recent surveys show a higher sex ratio for the cohort 
than earlier surveys.
The basic recall error pattern remains, with women married longer ago more
likely to report that their first-born is a boy.
Similarly, comparing women married in the same five-year period across surveys shows
that women married longer ago are more likely to report having a son.

@@

The relationship between length of marriage and recall error can also seen in 
Figures \ref{fig:sex_ratio_recall_rounds_bo1} and \ref{fig:sex_ratio_recall_rounds_bo2}, 
which show the observed sex ratio for children reported as first born as a function of 
duration of marriage at the time of the survey.
The solid line is the sex ratio of children reported as first born by the number of years 
between the survey and marriage, the dashed lines indicate the 95 percent confidence 
interval and the horizontal line the natural sex ratio (approximately 0.512).
To ensure sufficient cell sizes the years are grouped in twos.
In line with the results from Tables \ref{tab:recallMarriageBO1} and 
\ref{tab:recallMarriageBO2}, the observed ratio of boys is increasingly above the expected 
value the longer ago the parents were married.

@@

The increasingly unequal sex ratio with increasing marriage duration suggests that
a solution to the recall error problem is to drop observations for 
women who were married ``too far'' from the survey year.
The main problem is establishing what the best cut-off point should be, with the
trade-off between retaining enough observations and the correctness of the information.
As Tables \ref{tab:recallMarriageBO1} and \ref{tab:recallMarriageBO2} show, there are 
differences in recall error across the three surveys and between parities, although
this may simply be the result of differences in number of observations across surveys.
Furthermore, the recall error pattern is not entirely consistent across observed birth 
orders.
Since most of the surveys start showing significantly biased sex ratio from around 22
years of marriage on, I drop all observations where the marriage took place 22 years
or more.
